---
title: "Hierarchical Clustering"
author: "Shashidhar Reddy"
date: "2022-11-30"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
```{r}
#Displaying the required libraries
library(cluster)
library(caret)
library(dendextend)
library(knitr)
library(factoextra)
library(readr)
```

```{r}
#creating a data collection that solely includes numbers by importing a dataset
library(readr)
SB_Cereals <- read.csv("~/Downloads/Cereals.csv")
Num_data <- data.frame(SB_Cereals[,4:16])
```

```{r}
#Data lacking values should be removed
Num_data <- na.omit(Num_data)
```

```{r}
#Normalizing data
SB_Cereals_normalise <- scale(Num_data)
```

```{r}
#Use the normalized data to do hierarchical clustering using the Euclidean Dist technique.
Dist <- dist(SB_Cereals_normalise, method = "euclidean")
H_clust <- hclust(Dist, method = "complete")
```

```{r}
#the dendogram plotting process.
plot(H_clust, cex = 0.7, hang = -1)
```

```{r}
#Clustering with single linkage, full linkage, and the Agnes function,average linkage and Ward.
single_Hclust <- agnes(SB_Cereals_normalise, method = "single")
complete_Hclust <- agnes(SB_Cereals_normalise, method = "complete")
average_Hclust <- agnes(SB_Cereals_normalise, method = "average")
ward_Hclust <- agnes(SB_Cereals_normalise, method = "ward")
```

```{r}
#Choosing the most efficient course of action
print(single_Hclust$ac)
print(complete_Hclust$ac)
print(average_Hclust$ac)
print(ward_Hclust$ac)
```
#The ward strategy is the most successful one, as shown by its value of 0.9046042, which is evident given the facts provided.

# 2- Choosing the clusters:
```{r}
pltree(ward_Hclust, cex = 0.5, hang = -1, main = "Dendrogram of agnes (Using Ward)")
rect.hclust(ward_Hclust, k = 5, border = 2:7)
S_Group <- cutree(ward_Hclust, k=5)
D_frame_2 <- as.data.frame(cbind(SB_Cereals_normalise,S_Group))
```
```{r}
fviz_cluster(list(data = D_frame_2, cluster = S_Group))
```
#5 From the observation mentioned above, clusters can be selected.
#determining the stability and structure of the clusters. 

```{r}
#Building Partitions
set.seed(123)
One_partition <- Num_data[1:50,]
Two_partition <- Num_data[51:74,]
```

```{r}
#Performing Hierarchical Clustering while considering k = 5.
single_sb <- agnes(scale(One_partition), method = "single")
complete_sb <- agnes(scale(One_partition), method = "complete")
average_sb <- agnes(scale(One_partition), method = "average")
ward_sb <- agnes(scale(One_partition), method = "ward")
cbind(single=single_sb$ac , complete=complete_sb$ac , average= average_sb$ac , ward= ward_sb$ac)
pltree(ward_sb, cex = 0.6, hang = -1, main = "Dendogram of Agnes with Partitioned Data (Using Ward)")
rect.hclust(ward_sb, k = 5, border = 2:7)
cut_2 <- cutree(ward_sb, k = 5)
```

```{r}
#the centroids are calculated.
Sb_result <- as.data.frame(cbind(One_partition, cut_2))
Sb_result[Sb_result$cut_2==1,]
one_centroid <- colMeans(Sb_result[Sb_result$cut_2==1,])
Sb_result[Sb_result$cut_2==2,]
two_centroid <- colMeans(Sb_result[Sb_result$cut_2==2,])
Sb_result[Sb_result$cut_2==3,]
three_centroid <- colMeans(Sb_result[Sb_result$cut_2==3,])
Sb_result[Sb_result$cut_2==4,]
four_centroid <- colMeans(Sb_result[Sb_result$cut_2==4,])
centroids <- rbind(one_centroid, two_centroid, three_centroid, four_centroid)
x2 <- as.data.frame(rbind(centroids[,-14], Two_partition))
```

```{r}
#figuring out the Dist.
Dist_1 <- get_dist(x2)
Matrix_1 <- as.matrix(Dist_1)
dataframe1 <- data.frame(data=seq(1,nrow(Two_partition),1), Clusters = rep(0,nrow(Two_partition)))
for(i in 1:nrow(Two_partition)) 
  {dataframe1[i,2] <- which.min(Matrix_1[i+4, 1:4])}
dataframe1
cbind(D_frame_2$S_Group[51:74], dataframe1$Clusters)
table(D_frame_2$S_Group[51:74] == dataframe1$Clusters)
```
#12 of the above observations' findings are false, while 12 are true. We may thus claim that the model is only partially unstable.

# 3- The elementary public schools would like to choose a set of SB_Cereals to include in their daily cafeterias. Every day a different cereal is offered, but all SB_Cereals should support a healthy diet. For this goal, you are requested to find a cluster of “healthy Cereals''

```{r}
#Clustering Healthy SB_Cereals.
Healthy_SB_Cereals <- SB_Cereals
Healthy_SB_Cereals_RD <- na.omit(Healthy_SB_Cereals)
clust <- cbind(Healthy_SB_Cereals_RD, S_Group)
clust[clust$S_Group==1,]
clust[clust$S_Group==2,]
clust[clust$S_Group==3,]
clust[clust$S_Group==4,]
```

```{r}
#Mean ratings are used to select the best cluster.
mean(clust[clust$S_Group==1,"rating"])
mean(clust[clust$S_Group==2,"rating"])
mean(clust[clust$S_Group==3,"rating"])
mean(clust[clust$S_Group==4,"rating"])
```

#Cluster 1 may be chosen based on the data mentioned above because it is the highest.
#Therefore, Group 1 may be considered of as the cluster for a healthy diet.
